Capturing the geometric structure of our experiences and how we remember them

### 2018-07-24 main.tex
added text to results

To visualize the relationship between the video and recall event models, we embedded the video and average recall models into a 2D space (using the UMAP dimensionality reduction algorithm, for details see \ref{sec:methods}) where the points represent video/recall events and the distance between them represents their similarity in ``topic space'' (Fig.~\ref{fig:trajectory}a). (Note that all analyses are performed in the original 100 dimensional space.) We computed the group average transition probabilities from each event to each other event and plotted the values as lines where the tranparency is proportional to the probability of the transition (Fig.~\ref{fig:trajectory}b, see Supp Fig.~\ref{fig:ind_embeddings} for individual participant plots).  We observed that the two models have a very similar temporal evolution and geometric structure, and the transition probabilities tend to be strongest between events that occured sequentially (or nearby) in time.

@@

\begin{figure}[]
\centering
\includegraphics[width=1\textwidth]{figs/6_brains.pdf}
\caption{\small \textbf{Recall model searchlight analysis.} A). We used dynamic time warping to separately align each participant's recall model to the video model. The left matrix is S13's video-recall correlation matrix before warping. The yellow stars represent the path recovered by dynamic time warping. The right matrix is the result after warping. The same warping path was applied to the fMRI timeseries (separately for each participant). Below the matrices are toy examples representing a vectorized (aligned) recall correlation matrix and a vectorized (aligned) fMRI searchlight correlation matrix. These vectors were correlated for each searchlight sphere in the brain to assess brain regions whose temporal correlation structure mirrored the participant's recall model correlation structure. B) A surface plot of brain regions whose activity patterns mirrored the participant-specific recall correlation matrix. Significance threshold (p<.05) derived from a permutation procedure by phase shifting the video model and recomputing the z-scores.}
\label{fig:topics}
\end{figure}

>>>>

To visualize the relationship between the video and recall event models, we embedded the video and average recall models into a 2D space (using the UMAP dimensionality reduction algorithm, for details see \ref{sec:methods}) where the points represent video/recall events and the distance between them represents their similarity in ``topic space'' (Fig.~\ref{fig:trajectory}a).   We observed that the two models have a very similar temporal evolution and geometric structure. To further quantify this correspondence and to characterize how participants navigated through the space during recall, we analyzed the angle between successive recall events. We created a grid of evenly spaced points (.25 units) in the 2D ``topic space''. For each point on the grid, we drew a circle (radius=.25) around the point and grouped together all recall transitions that intersected the circle (across subjects). To visualize the average angle, we converted each transition angle to a unit vector and then averaged the vectors together. To assess consistency in the direction of the recall transition across participants, we performed a Rayleigh test (p<.001, corrected at p<.05 using permutation procedure described in \ref{sec:methods}). We found that on average, participants' recalls flowed along the same path as the video model (Fig.~\ref{fig:trajectory}b). Each participant's recall path can be seen in Fig.~\ref{fig:trajectory}c.

Notably, many of the participants' recall trajectories appear to recover the overall shape of the video stimulus. To quantify this similarity, we measured the average euclidean distance between each each segment of the video model (e.g. the line formed by successive video events) and the closest recall event for each participant.  Inuitively, if the participant approximately remembered every video event, the measure would approach 0. The more missing recall events, the higher the value. We found that this shape measure was highly correlated to the number of events remembered across participants (Pearson's $r$(16)=-.87, p$<$.001). While these measures are highly correlated (as might be expected), there are cases where the measures dissociate.  For example participant X and X recalled exactly the same number of events, but the shape measure is drastically different between them.  Conversely, participant X and X had the same shape measure, but there was a large difference in the number of events recalled. Thus, measuring similarity in shape between the video and recall models provides unique information about participant's recall behavior that is not captured by classic measures of memory accuracy.

@@

\subsubsection{Brain regions that mirror the correlational structure of memory}

\begin{figure}[]
\centering
\includegraphics[width=1\textwidth]{figs/6_brains.pdf}
\caption{\small \textbf{Recall model correlational structure searchlight analysis.} A). We used dynamic time warping to separately align each participant's recall model to the video model. The left matrix is S13's video-recall correlation matrix before warping. The yellow stars represent the path recovered by dynamic time warping. The right matrix is the result after warping. The same warping path was applied to the fMRI timeseries (separately for each participant). Below the matrices are toy examples representing a vectorized (aligned) recall correlation matrix and a vectorized (aligned) fMRI searchlight correlation matrix. These vectors were correlated for each searchlight sphere in the brain to assess brain regions whose temporal correlation structure mirrored the participant's recall model correlation structure. B) A surface plot of brain regions whose activity patterns mirrored the participant-specific recall correlation matrix. Significance threshold (p$<$.05) derived from a permutation procedure by phase shifting the video model and recomputing the z-scores.}
\label{fig:topics}
\end{figure}
